\chapter*{Preface}
\addcontentsline{toc}{chapter}{\protect\numberline{}Preface}

Well, well, well. This is something I wanted to write about for about two years now and haven’t had the time to actually sit down and write about it. I have been gathering this nuggets of knowledge for years and they have helped me to get a better look at myself, my relationships with my coworkers and the most importantly, how I actually do my job and how does this effect my work performance.

It’s been years since I was able to sit down in front of a TV and actually watch a movie without having minor panic attacks what is going on with the projects I was working on. Many of you probably know perfectly well what I am talking about. It’s this constant never ending pain inside that prohibits you of actually relaxing and have some quite time without brain rushing like you just drank 10 Red Bulls in an hour. The feeling of being on call all the time. Being available for everybody and not really having 5 minutes to slow down your thoughts. That feeling!

But regardless of all this I strive under pressure. I love it. I can’t live without it if I am being totally honest with myself. I love the grind. I love late night coding sessions. I love the subtle competition that comes with it. I just love it. And I find it impossible to explain this to my friends. They just can’t wrap their heads around it. How the hell can you enjoy this? How is this even remotely something compared to life? And to be honest, I can’t explain it to them. I have zero words. I just know that this is something I need in my life.

All this becomes problematic when pressure is not self inflicted but comes from the outside world. When other people start to immensely pressure you into dedicating time for them. Into agreeing with conditions you would never agree just so they leave you alone. And by the way, I find this type of extracting agreements from people totally disgusting. We all know that developers, when writing code, are in a totally different reality and if you ask them to give you their liver, they would probably say yes, just to be left alone with the joy of solving a problem. Taking advantage of this is just so foreign to me that it leaves me with a bad taste in my mouth. I hate it and I hate people that do this. They are intellectual and moral cowards as far as I am concern.

Dealing with this corporate bullshit for now 10 years really started to affect my better judgment. At one point I was working on three projects and sleeping for 3 hours a day for months on end. I slept behind a desk. I slept on the floor. I slept on tables. Practically everywhere but on a bed. And that didn’t bother me at all. I am that kind of a person that doesn’t need any special conditions in life. I don’t care about beds, about chairs, about tables, food, about day or night. I simply don’t care. When I start working on a project the only thing I care about is doing the work the best way that I can. I forget about all this stuff and this type of obsession can lead to real isolation. Which again, I just don’t mind.

And I am unapologetic about this. I don’t mind being called weird, strange, isolationist, odd, bizarre or any other adjective you might choose to describe me. I have my own morals and ethics that don’t revolve around external image. All I care about is internal happiness.

This book is not a set of rules. It’s not even something you or your company should blindly follow. Maybe you disagree with some of the points, or maybe even all of them completely. It’s fine, I don’t care actually. Maybe all this will do is make you think and reevaluate how you do your job.

And I am unapologetic about this. I don’t mind being called weird, strange, isolationist, odd, bizarre or any other adjective you might choose to describe me. I have my own morals and ethics that don’t revolve around external image. All I care about is internal happiness.

This book is not a set of rules. It’s not even something you or your company should blindly follow. Maybe you disagree with some of the points, or maybe even all of them completely. It’s fine, I don’t care actually. Maybe all this will do is make you think and reevaluate how you do your job.
